Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-05 foreground_delay_spectrum_modeling.tex
From v0.3.2: ===========
(*) Created branch v0.3.2b
(*) Removed some of the comments

 Detection and measurement of 21cm HI emission from the Epoch of Reionization, at redshifts above 6, is limited primarily by foreground emission. With much of the information coming from the spectral line that maps to time and line of site distance, foregrounds affecting the spectral domain are deserving of close attention, particularly those that are due to the instrument because they are in theory removable.  Here we investigate in detail the chromatic appearance of flat spectrum foregrounds imprinted by the instrumental transfer function. Using the delay spectrum technique to measure power spectrum contamination on a per-baseline basis we demonstrate that the foreground signatures which have the largest impact on the HI signal arise from power entering far from the center of the field of view.  Comparing data from recent Murchison Widefield Array observations with simulations broken down by sky position and type of emission, we identify smooth emission near the horizon as the largest contributing factor, followed by bright sources in the same locations. The observation that these contaminants only effect a fraction of the baselines leads us to propose that by down-weighting certain baselines at certain times, a large fraction of foreground contamination could be mitigated.

@@

Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness\citep{mor06,bow09,liu11,par12,dil13}.  Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.  A general model of how smooth spectrum foregrounds are scattered to higher $k$-modes by the wide-field response of the instrument, into the so-called `wedge',  has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to throw out $k$-modes which could be contaminated (Avoidance).  The more aggressive option is to subtract a sky model and regain access to modes lost by Avoidance. In both cases one large unknown is which parts of the sky are most critical to either Avoid or Subtract. Here we will focus primarily on the Avoidance strategy, asking which foreground components are most likely to ``leak'' from foreground modes to Reionization modes.

Foregrounds work their way into the power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in the estimating the power spectrum. Here we are focusing on the instrumental effects. To de-couple from possible analysis effects, such as those pointed out by \citet{Haz13}, we compute power spectra using the per-baseline approach of \citet{pob13,arp14}.  This approximates the power spectrum as the fourier transform of the spectra output by the correlator.

 Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline.  We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon. We also find that the smooth galactic emission plays a larger role then previously thought. We quantify this contamination by breaking the simulations down by sky region and type of emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to downweight badly contaminated baselines.

In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, \S\ref{sec:obsparms}, and \S\ref{sec:data-analysis} we describe the MWA setup, summarize the observing parameters and present the resulting data.  Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.  \S\ref{sec:discussion} offers a discussion along with a basic comparison to other current and future instruments. \S\ref{sec:conclusion} offers conclusions and an initial description of a more precise foreground avoidance technique.

@@

In order to make a quantitiative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$0.7 and $\sim$0.3 respectively. In addition, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise. We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively. % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.

% Figure~\ref{fig:data-sim-ratio} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}. The probability densities and cumulative distributions are shown in black and gray respectively. Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$. The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

@@

% The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors. Due to numerous combinations of such factors, we use unique codes for these combinations as described in table~\ref{tab:codes}. The first column refers to the place holder of the specific parameter in the code sequence from left to right. The parameter codes are specified in this order. The second column gives the parameter used in describing features in delay spectrum. The third column denotes the code describing the parameter. Snapshot index uniquely specifies the snapshot and carries a numeric code (1 and 2 for off-zenith and on-zenith snapshots respectively) leading the code sequence. Following this, the type of emission is specified as compact, diffuse, Galactic center, or Galactic plane using codes C, D, GC, or GP respectively. Although in reality, a feature could be due to a combination of emission types, the code we use refers to the predominant type of emission that can be attributed to the feature. The next code specifies the direction of this predominant source of emission. The code values for the direction of emission are the usual abbreviations of the cardinal and ordinal directions besides Z (zenith). Following this is the code for the baseline orientation bin in the half-plane specified as S, SE, E, and NE along directions south, south-east, east, and north-east respectively. Code `A' is used to refer to all baselines. Finally, the parameter describing the region of antenna power pattern coincident with the predominant emission responsible for the feature is specified as P (primary lobe) and S (secondary lobe). The secondary lobe code carries a numeral suffix to further specify uniquely the sidelobe in the direction of emission. For example, a feature code 1-D-W-E-P indicates it is from the first snapshot (1) arising due to diffuse emission (D) westward (W) of zenith observed by predominantly eastward (E) baselines centered on the primary beam (P) of the antenna power pattern. Simiarly, feature code 2-GC-W-NE-S3 implies the feature is seen in the second snapshot (2) due to the Galactic Center (GC) that is westward (W) of zenith by north-eastward (NE) baselines and the region coincides with the third sidelobe (S3) in the direction of emission.

% \begin{table*}[htp]
% \caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
% \centering
% \begin{tabular}{ccc}
% \tableline\tableline\tableline
% Order & Parameter Description & Code \\
% \tableline\tableline
% 1 & Snapshot index & 1, 2, \ldots \\
% 2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
% 3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
% & & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
% 4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
% 5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
% & (in the direction of emission) & \\
% \tableline
% \end{tabular}
% \end{table*}

@@

% Figure~\ref{fig:dsm} shows the noiseless delay spectra due to diffuse emission. Figure~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} correspond to off-zenith and on-zenith snapshots respectively. In general, shorter baselines are expected to be receptive to large scales such as from diffuse emission as well as from small scales from compact objects. This effect is most clearly visible in the on-zenith snapshot noted by feature code 2-D-Z-A-P which stands for the signature observed due to diffuse emission from zenith direction on short baselines of all orientations restricted primarily to the primary beam. This is a notable feature on baselines of length $\lesssim 60$~m. For a frequency of 185~MHz, this length scale corresponds to angular scales of $\gtrsim 1\fdg 5$ on the sky. This feature is noted only for diffuse emission from inside the primary beam.

% The most prominent feature, however, is 1-GC-W-E-S3 in figure~\ref{fig:dsm_delay_spectrum_snap0}. This is noted in the off-zenith snapshot caused by the bright Galactic center which is on the west at an altitude of $\sim 33$\arcdeg with $\sim 100$~minutes remaining before it sets. This emission is observed as it is coincident with the third sidelobe of the power pattern in the westward direction with a neagtive delay of value almost near the horizon limit on eastward baselines.

% \begin{figure*}[htp]
% \centering
% \subfloat[][Off-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
% \subfloat[][On-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
% \caption{Delay spectrum visibility amplitudes (in units of Jy~Hz) from diffuse foreground emission obtained on baselines of length $\lesssim 314.7$~m. The left and right panels correspond to off-zenith and on-zenith snapshots respectively. These are noiseless delay spectra obtained after {\it CLEAN} deconvolution along delay axis. A logarithmic color scale is used. Most prominent features are due to the bright Galactic center shining through the third sidelobe (1-GC-W-E-S3), diffuse emission of short baselines at zero delays (2-D-Z-A-P), and diffuse emission on long baselines near the horizon limit leaking in through the second sidelobes (1-GP-NE-NE-S2, 2-GP-N-N-S2, and 2-GP-S-N-S2). Features from diffuse emission are predominantly edge-heavy resembling a two-pronged fork. See text for a detailed description of these features.}
% \label{fig:dsm}
% \end{figure*}

@@

% Interestingly, in contrast to feature code 2-D-Z-A-P, all the other features in figures~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} extend from the shortest ($\sim 7.7$~m) to longest baselines ($\sim 314.7$~m). We argue this is because even wider baselines are shortened when projected against directions along the baseline vector towards the horizon while considering an all-sky foreground model. Such a shortening due to projection effects allows diffuse emission from near the horizon to be observed at maximum delays that correspond to the horizon limit even on long baselines. Feature 1-GP-NE-NE-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0}, and features 2-GP-N-N-S2 and 2-GP-S-N-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0} demonstrate this effect of shortening of baselines. 1-GP-NE-NE-S2 corresponds to emission observed from the Galactic plane from north-eastward directions by north-eastward baselines from the second sidelobe in their power pattern. Features 2-GP-N-N-S2 and 2-GP-S-N-S2 are also Galactic plane emission from southward and northward directions observed by second sidelobes of northward baselines with opposite delays. Owing to their apparent structure in delay spectra, we hereafter refer broadly to this edge-heavy structure as the ``two-pronged fork'' feature.

% Thus, the most notable features from diffuse emission are those from primary beam area on short baselines, bright Galactic center from the third sidelobe, and the two-pronged fork.

@@

% Figure~\ref{fig:csm} shows the noiseless delay spectra obtained from compact foreground objects. The off-zenith and on-zenith snapshot delay spectra are shown in figures~\ref{fig:csm_delay_spectrum_snap0} and \ref{fig:csm_delay_spectrum_snap1} respectively.

% In the off-zenith snapshot, the most prominent feature 1-C-E-E-P arises due to eastward baselines observing the eastward sky at positive delays centered on the primary beam. The next prominent feature 1-C-E-N-P at zero delays arises due to the compact emission from eastward sky centered on the primary beam observed by northward baselines. Features 1-C-NE-NE-S2 and 1-C-W-E-S3 at positive and negative horizon limits are caused by north-eastward sky emission observed by the second sidelobe of north-eastward baselines and westward sky centered on the third sidelobe of eastward baselines respectively.

% The most prominent feature in the on-zenith snapshot is 2-C-Z-A-P appearing at zero delays caused by compact sources at zenith observed by all baselines along their primary beams. First and second sidelobes along north and south observed by northward baselines contribute to symmetric positive and negative delays as demonstrated by features 2-C-S-N-S1, 2-C-N-N-S1, 2-C-S-N-S2, and 2-C-N-N-S2.

% \begin{figure*}[htp]
% \centering
% \subfloat[][Off-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
% \subfloat[][On-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
% \caption{Same as figure~\ref{fig:dsm} but for compact foreground objects. Color scale is the same as in figure~\ref{fig:dsm}. The different features annotated are a result of emission observed by compact foreground objects through primary beam and sidelobes. Broadly, signatures in delay spectra of compact foreground objects are center-heavy in appearance. See text for a detailed description of the different features.}
% \label{fig:csm}
% \end{figure*}

% While compact foreground objects give rise to features throughout the foreground wedge, the primary distinction between delay spectra of compact and diffuse foregrounds is that the latter are predominantly center-heavy when compared to edge-heavy features of the former.

% % Figure~\ref{fig:obs-begin} shows the delay spectra obtained with the compact foreground model described in \S\ref{sec:CSM} at the beginning of the observation when the MWA tiles are phased to point at RA = 4\fdg 387, Dec = -29\fdg 86. The contours of the antenna power pattern are shown in the central panel. The flux densities of compact foreground objects multiplied by the antenna power pattern at the respective locations are shown on a logarithmic color scale right below the central panel. For reference, maps of delays between the received signals are shown for 100~m baselines with $\theta_b=0\arcdeg, 45\arcdeg, 90\arcdeg, \textrm{and}\,135\arcdeg$ to relate the strength of emission with the corresponding delay in the delay spectra.

% % \begin{figure*}[htb]
% % \centering
% % % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_0.eps}
% % \caption{Delay spectra, compact foreground sources, and delay maps for a snapshot at the beginning of the observation pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at LST = 22.08 hours. The phase center is at RA = -28\fdg 8, Dec = -26\fdg 701. The baselines have been binned by their orientation ($\theta_\textrm{b}$) anti-clockwise from east. The center-right panel shows delay spectra from baselines oriented between -22\fdg 5 $\le \theta_\textrm{b}<$+22\fdg 5. The upper-right, upper-center, and upper-left panels have baseline orientations lying between +22\fdg 5 $\le \theta_\textrm{b}<$+67\fdg 5, +67\fdg 5 $\le \theta_\textrm{b}<$+112\fdg 5, and +112\fdg 5 $\le \theta_\textrm{b}<$+157\fdg 5, respectively. The color scale in these delay spectra are given by the color bar on top. The panels diametrically opposite to aforementioned panels represent the delay map on the sky for a 100~m baseline oriented along east, north-east, north, and north-west respectively. The color scale for these delay maps is specified at the bottom. Delays are positive towards east and north sky positions for eastward and northward baseline orientations respectively. The central panel shows the compact foreground model multiplied by the antenna power pattern. The color scale is specified below the panel. Also shown are contours of the antenna power pattern. The contour levels shown are: 0.0078125, 0.03125, 0.125, and 0.5.}
% % \label{fig:obs-begin}
% % \end{figure*}

% % \begin{figure*}[htb]
% % \centering
% % % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_1.eps}
% % \caption{Same as figure~\ref{fig:obs-begin} but for a snapshot during the middle of the observation pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at LST = 0.09 hours. The phase center is at RA = 1\fdg 35, Dec = -26\fdg 701. }
% % \label{fig:obs-middle}
% % \end{figure*}

@@

% Figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} show the delay spectra for the composite all-sky foreground model. While features from both its constituents occupy and overlap in different regions of the foreground wedge, their relatives magnitudes determine the resultant signatures seen in the delay spectra. For instance, the fork-shaped features of diffuse emission are brighter by a factor $\gtrsim 10$ relative to the compact foreground features at the edges of delay spectra in the respective snapshots. Hence, features 1-C-W-E-S3 and 1-C-NE-NE-S2 in figure~\ref{fig:csm_delay_spectrum_snap0} have been masked by features 1-GC-W-E-S3 and 1-GP-NE-NE-S2 respecively from diffuse emission in figure~\ref{fig:dsm_delay_spectrum_snap0}. Similarly, the edge features 2-C-S-N-S2 and 2-C-N-N-S2 from compact objects in figure~\ref{fig:csm_delay_spectrum_snap1} have been masked by diffuse emission features 2-GP-S-N-S2 and 2-GP-N-N-S2 respectively in figure~\ref{fig:dsm_delay_spectrum_snap1}.

% On the other hand, in central and inner regions of the wedge, clearly the center-heavy compact foreground model features dominate by $\gtrsim 2$ orders of magnitude relative to features arising out of diffuse emission. This is quite evident since almost all the diffuse emission features are confined to the edges of the wedge. Further, 2-D-Z-A-P in figure~\ref{fig:dsm_delay_spectrum_snap1} is completely masked by 2-C-Z-A-P in figure~\ref{fig:csm_delay_spectrum_snap1}. Thus the inner regions of the wedge consist only of compact foreground features 1-C-E-N-P and 1-C-E-E-P in figure~\ref{fig:asm_delay_spectrum_snap0} and 2-C-Z-A-P, 2-C-S-N-S1 and 2-C-N-N-S1 in figure~\ref{fig:asm_delay_spectrum_snap1}.

% Figures~\ref{fig:baseline_composite_asm_delay_spectrum_snap0} and \ref{fig:baseline_composite_asm_delay_spectrum_snap1} show the delay spectra from the diffuse and compact foreground models as well as from the composite model for the off-zenith and on-zenith snapshots respectively for a selected baseline labeled as ``155-154''. This baseline is of length $\simeq 201$~m and orientation $\theta_b\simeq 13\fdg 7$. These are effectively vertical slices of figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} through the baseline location. The horizon delay limits are shown by vertical gray lines. Diffuse emission is negligible in the off-zenith snapshot except at the negative horizon delay limit. This is identified to be feature 1-GC-W-E-S3 noted in figure~\ref{fig:asm_delay_spectrum_snap0} due to the bright Galactic center on the westward sky observed by eastward baseline through its third sidelobe. All the other features are predominantly due to compact foreground objects. The peak at delay of $\sim 0.3\,\mu$~s is identified as feature 1-C-E-E-P in figure~\ref{fig:asm_delay_spectrum_snap0}. The peak at zero delay is from feature 1-C-Z-A-S1 while the other secondary peaks arise compact sources coincident with sidelobes of the power pattern. In the on-zenith snapshot, the feature 1-C-E-E-P has transformed to feature 2-C-Z-A-P due to change in pointing position. The diffuse emission is confined to the edges labeled by features 2-D-W-E-S2 and 2-D-E-E-S2. All intermediate peaks are due to compact foreground objects observed through the different sidelobes of the antenna power pattern.

% \begin{figure*}[htp]
% \centering
% \subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
% \subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
% \subfloat[][Off-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_0.edited.eps}}
% \subfloat[][On-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_1.edited.eps}}\\
% \caption{{\it Top}: Same as figure~\ref{fig:dsm} but for a composite all-sky foreground model consisting of both diffuse emission and compact objects. The color scale used is the same. The different features annotated arise from those in figures~\ref{fig:dsm} and \ref{fig:csm}. Inner regions of the wedge are dominated by center-heavy features such as 1-C-E-N-P, 1-C-E-E-P, 2-C-Z-A-P, 2-C-N-N-S1 and 2-C-S-N-S1 characteristic of compact objects. Outer regions are dominated by edge-heavy fork-shaped features such as 1-GC-W-E-S3, 1-GP-NE-NE-S2, 2-GP-S-N-S2 and 2-GP-N-N-S2 characteristic of diffuse emission. The resultant structure in delay spectrum takes the form of a {\it pitchfork}. {\it Bottom}: Slices of delay spectrum visibility amplitudes from the top panel at the location of a selected eastward baseline. The baseline vector is $\sim 201$~m in length oriented $\sim 13\fdg 7$ measured anti-clockwise from east. The foreground components are: diffuse sky model (DSM, dashed lines), sky made of compact sources (CSM, dotted lines) and all-sky composite model (ASM, solid lines). The gray vertical lines signify the wedge boundaries set by the horizon delay limit. Compact sources dominate the delay spectra in the inner part of the wedge as marked by features 1-C-E-E-P and 2-C-Z-A-P. Edge-heavy fork-shaped features such as 1-GC-W-E-S3, 2-D-W-E-S2 and 2-D-E-E-S2 from diffuse emission are more significant at the edges of the wedge. These give rise to a ``pitchfork'' appearance. Other secondary peaks inside the wedge are due to compact objects observed through sidelobes of the antenna power pattern. The secondary bumps outside the wedge in all the model components is a result of imperfect deconvolution along delay axis. Refer to text for detailed description of these features.}
% \label{fig:asm}
% \end{figure*}

% Thus the composite all-sky foreground model results in a delay spectrum structure that captures both the edge-heavy two-pronged fork structure from diffuse foreground emission and the center-heavy features from compact foreground objects. Hereafter, we refer to this structure as a ``three-pronged pitchfork''.

>>>>

We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using an antenna pair based ``delay spectrum'' analysis technique. We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies. % Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution. 
We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse and compact Galactic and extragalactic emission. By tying model parameters to those of a subset of first season of observing with the MWA, our simulations agree with the observed data to within expected uncertainty. A wedge--shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed. Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge--heavy {\it two--pronged fork}--like features even on wide antenna spacings, hitherto not predicted. Features in the inner regions of the wedge are predominantly due to compact foreground objects. This results in a {\it pitchfork}--shaped delay spectrum. The bright Galactic center severly contaminates the {\it EoR window} on certain antenna spacings in certain observations. For such cases, we present an effective practical tool for screening interferometers with high foreground contamination and mitigating their effects on EoR data analysis.

@@

One of the key challenges in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which lack such symmetry. Foreground removal techniques rely on the spectral smoothness of foreground emission \citep{mor06,bow09,liu11,par12,dil13}.

Precise understanding of these foreground signatures is necessary in order to remove foreground systematics from the measured power spectrum. Considerable effort has been made in mapping the residual foreground signatures in observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} after foreground subtraction.

In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum. The simplicity of this technique and usage of basic interferometric visibility data fourier transformed along frequency (delay spectrum) on individual interferometers offer very effective diagnostics for scheduling observations and data analysis, and an insightful view onto foreground signatures in the EoR H{\sc i} power spectrum. We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA. This simple technique allows us to predict foreground signatures, hitherto not predicted.

We have organized this paper as follows. \S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum. \S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively. Analysis of data obtained with the MWA is described in \S\ref{sec:data-analysis}. In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the models used for the antenna power pattern and foreground. Here, we also establish that the results of modeling and data analysis agree with each other. In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters. In \S\ref{sec:fg-grading}, we use our understanding to grade interferometers based on the amount of foreground contamination estimated by the delay spectrum technique. In \S\ref{sec:summary}, we present our summary and conclusions.

@@

In order to make a quantitiative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$0.7 and $\sim$0.3 respectively. In addition, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise. We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
\end{figure}

@@

@@

@@

@@

@@
